---
title: "MLR"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: spacelab
    code_folding: hide
editor_options: 
  chunk_output_type: console
---
```{r setup, include=FALSE}
library(tidyverse) ## Data Manipulaion
library(patchwork)
library(readxl)
library(dplyr)


knitr::opts_chunk$set(
  fig.width = 8,
  fig.asp = 1,
  out.width = "90%",
  echo = TRUE,
  message = FALSE)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis")

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

After visualization of the correlation of happiness score and other factors in different levels, we decide to formally test the main effect of `gdp`, `social support`, `life_expectancy`, `freedom`, `positive_affect`, `negative_affect` , `generosity` and `corruption` on the happiness score and the `interaction` effect between several covariates. 

By the inspiration of exploratory correlation plots, the developing levels and continent may also influence the estimates in the regression model. However, wwe think that the difference in correlation coefficient between continents mainly due to the developing level. . Besides, we define the developing level based on the gdp, so we exclude gdp and only include the developing level as a categorical variable in the final analysis. 

* After fitting with total of 8 main effects, only five of them showing significant results (P<0.001), so we only use `social support`, `life_expectancy`, `positive_affect` and `corruption` to interact with `develop`. 

```{r}
happy = read_csv("data/happy.csv")
data =  happy %>% 
  mutate(develop = as.factor(develop),
         continent = as.factor(continent)) %>%
  select(-c(country, year, o_gdp))

fit1 = lm(happiness_score~ 
            freedom + 
            negative_affect + 
            generosity + 
            social_support*develop +  
            life_expectance*develop + 
            positive_affect*develop + 
            corruption*develop, data = data)

```

* There is a significant interaction effect of `social_support` and `develop`, `corruption` and `develop`. The effect of `social_support` and `corruption`  on the Happiness Score is depending on the value of `develop` at 0.001 significance level.

* There is no overall significant effect of either `develop` or `life_expectance`, but there is a significant crossover interaction at 0.001 significance level. The effect of `life_expectance` on the Happiness Score is opposite, depending on the value of `develop`.

To interpret the results, we conduct the stratified analysis based on the developing level. Results are below, and model assumptions are checked after the discussion.

```{r}
developing =  happy %>% 
  filter(develop == "developing") %>% 
  select(-c(o_gdp, develop, continent, country, year)) 
  
developed =  happy %>% 
  filter(develop == "developed") %>% 
  select(-c(o_gdp, develop, continent, country, year)) 

fit2 = lm(happiness_score~., data = developing)
fit3 = lm(happiness_score~., data = developed)

```

* For developing countries, there is a significant positive effect of  `gdp`, `social_support`, `life_expectance` and `positive_affect` on the Happiness Score, at 0.001 significance level. 

```{r}
fit2 %>% broom::tidy() %>% knitr::kable(digits = 3)
```

* For developed country, there is significant positive effect of  `social_support`, `positive_affect` and `generosity`, and there is significant negative effect of `corruption` on the Happiness Score, at 0.001 siginificance level. 

```{r}
fit3 %>% broom::tidy() %>% knitr::kable(digits = 3)
```


## Discussion

Although the happiness score slightly increases from 2007 to 2018 in the worldwide level, some regions have decreasing trend, and developed countries also have decreasing trend. It is interesting that the happiness score does not increase given the economic growth.

From the correlation analysis and the stratified analysis in linear regression, the happiness score is highly associated with the `social support` and `positive affect` both in developed countries and developing countries. However, `gdp` and `life expectancy` are positively associated with happiness score only in the developing countries, while `generosity` is positively , and `corruption` is negatively associated with happiness score only in the developed countries. 

These results are reasonable and mostly what we expect. Overall, happiness score depends on economics (GDP), social support (someone to count on), health (life expectancy), positive mood (positive affect) and perceptions of corruption. For developing countries, economics and health become more important in evaluating happiness, while for developed countries, perceptions of corruption and generosity(donation) become more crucial. 

The government of developing countries should focus more on economic growth and improving healthcare to increase the happiness perception. The government of developed countries should have policies to control the corruption in the  business and government and have policies to encourage donation to increase the happiness perception. 




