---
title: "Exploratory Analysis"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: spacelab
    code_folding: hide
editor_options: 
  chunk_output_type: console
---
```{r setup, include=FALSE}
library(tidyverse) ## Data Manipulaion
library(corrplot) ## Correlation plot
library(patchwork)
library(readxl)
library(dplyr)
library(countrycode)  # Gets country code 
library(gganimate)
library(ggplot2)


knitr::opts_chunk$set(
  fig.width = 8,
  fig.asp = 1,
  out.width = "90%",
  echo = TRUE,
  message = FALSE)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis")

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

To better show the association of other factors and happiness score, we standardize the value of other factors by subtracting mean and dividing standard error. This is because  the scale of `gdp`  and  `life_expectancy` are dramatically different from other factors. 

```{r Standardaization, warning=FALSE}
happy =  read_csv("data/happy.csv")
happy_standard = happy %>% 
  mutate(s_gdp = scale(gdp),
         s_social_support = scale(social_support),
         s_life_expectance = scale(life_expectance),
         s_freedom = scale(freedom),
         s_positive_affect = scale(positive_affect),
         s_negative_affect = scale(negative_affect),
         s_generosity  = scale(generosity),
         s_corruption  = scale(corruption)) %>%
  select(-c(gdp:corruption), -o_gdp)
```

## Exploratory analysis
### **Happiness score of all countries changes over time** 

* Worldwide
```{r, warning=FALSE, out.width = "70%"}
plot1 = happy %>% 
  group_by(year) %>% 
  summarise(year_mean_score = mean(happiness_score)) %>% 
  ggplot(aes(x = year, y = year_mean_score)) +
  geom_path() +
  labs(title = "Happiness score changes worldwide: 2007-2018",
       x = "Year",
       y =  "Happiness score") +
  scale_x_continuous(
    breaks = c(2007,2008,2009,2010,2011, 2012, 2013,2014,2015,2016,2017,2018))

plot1+transition_reveal(year)
```


From the first plot, we can easily find that the mean happiness score increases from 2007 to 2010, and gradually decreases to the lowest point in 2014. Finally, it increases in recent years. 

* By continent and developing level
```{r, warning=FALSE}
plot2 = happy %>% 
  group_by(year, continent) %>% 
  summarise(year_mean_score = mean(happiness_score)) %>% 
  ggplot(aes(x = year, y = year_mean_score, color = continent)) +
  geom_point() +
  geom_line() +
  geom_smooth(method = lm, size  = 0.5, 
              se = FALSE, color = "red") +  
  facet_grid(. ~ continent) +
  labs(title = "Happiness score changes in various continent: 2007-2018",
       x = "Year",
       y =  "Happiness score") +
 theme(legend.position = "none")
plot3 = happy %>% 
  drop_na(develop) %>%
  group_by(year, develop) %>% 
  summarise(year_mean_score = mean(happiness_score)) %>% 
  ggplot(aes(x = year, y = year_mean_score, color = develop)) +
  geom_point() +
  geom_line() +
  geom_smooth(method = lm, size  = 0.5, 
              se = FALSE, color = "red") +  
  facet_grid(. ~ develop) +
  labs(title = "Happiness score changes in developed and developing countries: 2007-2018",
       x = "Year",
       y =  "Happiness score")+
 theme(legend.position = "none")

plot2/plot3

```

From the second plot, we divide these countries into different regions by the geographic continent. Australia has the overall highest and Africa has the overall lowest happiness score. Only Europe and Asia show the general increasing trend of happiness score over time. 

We can roughly further divide these continents into three groups base on the happiness scores. The high happiness score group contains Australis, the medium happiness score group contains Europe, North America and South America, and the low happiness score group contains Africa and Asia. 


From the third plot, the mean happiness score is obviously higher in developed countries than it in developing countries. 


### **The correlation plot for happiness score and other factors: worldwide**


```{r, warning=FALSE, out.width = "70%",fig.align="center"}
happy_standard %>% 
  pivot_longer(s_gdp:s_corruption,
               names_to = "standard_factors",
               values_to = "standard_value") %>% 
  ggplot(aes(x = happiness_score, y = standard_value,
             color  = standard_factors)) +
  geom_point() +
  geom_boxplot(alpha = .8) +  
  geom_smooth(method = lm, color = "red") +
  facet_grid(. ~ standard_factors) +
  labs(x = "Happiness score", 
       y = "Standard score of other factors", 
       title = "Happiness score against other factors worldwide: 2007-2018") + 
  theme(legend.position = "none")

happy %>% 
  drop_na() %>%
  select(-c(country, year, continent, develop,o_gdp)) %>% 
  cor() %>%
  corrplot.mixed(title = "Happiness score against other factors worldwide: 2007-2018",
                 tl.cex = 1.2, tl.pos = "lt", number.cex = 1.2,
                  mar = c(0,0,1,0))
```


* From the scatterplot, the factors `negative_affect` and `corruption` have the negative association with happiness score, but `gdp`, `social support`, `life_expectancy`, `freedom`, `positive_affect` and `generosity` have the positive association.

* From the correlation matrix plot, the happiness score is highly associated with `gdp`, `social_support` and `life_expectancy`(> 0.7). In addition, `gdp` and `life_expctancy` are highly correlated with each other(> 0.7).

### **The correlation plot for happiness score and other factors : continent**

```{r, warning=FALSE, out.width = "70%",fig.align="center"}
happy %>% 
  drop_na() %>%
  filter(continent == "Australia") %>% 
  select(-c(country, year, continent, develop,o_gdp)) %>% 
  cor() %>%
  corrplot.mixed(title = "Australia: 2007-2018",
                 tl.cex = 1.2, tl.pos = "lt", number.cex = 1.2,
                  mar=c(0,0,1,0))

happy %>% 
  drop_na() %>%
  filter(continent %in% c("Europe", "North America", "South America"))  %>% 
  select(-c(country, year, continent, develop,o_gdp)) %>% 
  cor() %>%
  corrplot.mixed(title = "Europe, North America, South America: 2007-2018",
                 tl.cex = 1.2, tl.pos = "lt", number.cex = 1.2,
                  mar = c(0,0,1,0))

happy %>% 
  drop_na() %>%
 filter(continent %in% c("Africa", "Asia")) %>% 
  select(-c(country, year, continent, develop,o_gdp)) %>% 
  cor() %>%
  corrplot.mixed(title = "Africa, Asia: 2007-2018 ",
                 tl.cex = 1.2, tl.pos = "lt", number.cex = 1.2,
                  mar=c(0,0,1,0))
```

* Australia has the highest happiness score, and the score does not have a high correlation with other factors(not greater than 0.4). `gdp` is strongly positively associated with `corruption`(0.69) and negatively associated with `positive_affe t`(-0.86) 

* Europe, North America, South America  have the medium happiness score, and the score has a moderate correlation with `gdp`, `social support`, `life_expectancy`, `freedom`, `positive_affect` and `corruption`(0.6 - 0.7) 

* Africa and Asia have the lowest happiness score,  and the score has a moderate correlation with `gdp`, `social support` and `life_expectancy`(0.6 - 0.7).


### **The correlation plot for happiness score and othe factors : developing level**

```{r,  warning=FALSE, out.width = "70%",fig.align="center"}
happy %>% 
  drop_na() %>%
  filter(develop == "developing") %>% 
  select(-c( country, year, continent, develop,o_gdp)) %>% 
  cor() %>%
  corrplot.mixed(title = "Developing: 2007-2018",
                 tl.cex = 1.2, tl.pos = "lt", number.cex = 1.2,
                  mar=c(0,0,1,0))

happy %>% 
  drop_na() %>%
  filter(develop == "developed") %>% 
  select(-c(country, year, continent, develop,o_gdp)) %>% 
  cor() %>%
  corrplot.mixed(title = "Developed: 2007-2018",
                  tl.cex = 1.2, tl.pos = "lt", number.cex = 1.2,
                  mar=c(0,0,1,0))
```

* Developing countries have lower happiness score, and the score has a moderate correlation with `gdp`, `social support` and `life_expectancy`(around  0.6). 

* Developed countries have higher happiness score , and the score has a moderate correlation with `social support`,`freedom`, `positive_affect`, `negative_affect`, `generosity` and `corruption` (around 0.6)


